Paper 1 Notebook

### 2023-12-12 minos/outcomes/paper_1_notebook.Rmd
adding some mice analysis

# baseline OLS model formula. 
raw_model_formula_string <- "SF_12 ~ factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(labour_state_raw) + 
                  relevel(factor(NSSEC), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_netinc)"

@@

# get only needed variables and replace missing codes with NA.
raw.data.colnames <- colnames(raw.data)
raw.data.colnames[which( colnames(raw.data)=="job_sec" )] <- "NSSEC"
colnames(raw.data) <- raw.data.colnames
raw.data <- raw.data[, append(variables, c("time", 'pidp', 'weight'))]
raw.data <- replace.missing(raw.data)

@@

# get raw data file names.
full.raw.filelist <- list.files(raw.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[21:30] # get files for 2011-2013.
# load raw data.
full.raw.data <- do.call(rbind, lapply(full.raw.filelist, read.csv))
full.raw.data.colnames <- colnames(full.raw.data)
full.raw.data.colnames[which( colnames(full.raw.data)=="job_sec" )] <- "NSSEC"
colnames(full.raw.data) <- full.raw.data.colnames
# get only needed variables and replace missing codes with NA.
full.raw.data <- full.raw.data[, append(variables, c("time", 'pidp', 'weight'))]
full.raw.data <- replace.missing(full.raw.data)
```

@@

# calculate baseline OLS model.
sf12.lm <- lm(raw_model_formula_string,
              data= raw.data)
# print summary statistics
print(summary(sf12.lm))

@@

start.data <- raw.data[which(raw.data$time==2013),]
handover.data <- start.data[, c("pidp", "SF_12", "time")]
for (i in 1:7){
  start.data$SF_12 <- predict(sf12.lm, start.data)
  start.data[, 'time'] <- start.data[, 'time'] +1
  handover.data <- rbind(handover.data, start.data[, c("pidp", "SF_12", "time")])
}

@@

mice.data.colnames <- colnames(mice.data)
mice.data.colnames[which(colnames(mice.data)=="job_sec" )] <- "NSSEC"
colnames(mice.data) <- mice.data.colnames

@@

mice.2018.pidp <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "pidp"]
mice.2018.time <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "time"]
mice.2018.weights <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "weight"]
```

@@

imp_columns <- c("SF_12",
                 'sex',
                 "NSSEC",
                 "education_state",
                 "S7_labour_state", 
                 "region", 
                 "ethnicity", 
                 "age", 
                 "time", 
                 "hh_income",
                 "housing_quality",
                 "neighbourhood_safety",
                 "loneliness",
                 "ncigs",
                 'nutrition_quality'
)

@@

aggr_columns <- c("SF_12",
                 "NSSEC",
                 "education_state",
                 "S7_labour_state", 
                 "ethnicity", 
                 "hh_income",
                 "neighbourhood_safety",
                 "loneliness",
                 "ncigs",
                 'nutrition_quality'
)
# initial missingness structure for 2011-2013 US data.
pdf(paste0(save.path, "total_missingness_structure1.pdf"))
aggr(mice.data[,aggr_columns], sortVars=T, prop=T, combined=F,varheight=T, oma=c(10,5,1,2), gap=0)
dev.off()

mice_columns <- c("age", 
                  "region", 
                  #"heating", 
                  "NSSEC", 
                  "ncigs",
                  "education_state",            
                  "ethnicity",
                  "loneliness",
                  "sex", 
                  "SF_12",
                  #"SF_12p",
                  #"smoker",
                  "nkids",       
                  "behind_on_bills",
                  "financial_situation",
                  "future_financial_situation",
                  "likely_move",
                  "ghq_depression",
                  "ghq_happiness",
                  "clinical_depression", 
                  "scsf1",
                  "health_limits_social",
                  #"hhsize",
                  #"housing_tenure",
                  #"urban", 
                  "housing_quality",
                  "hh_income",
                  "neighbourhood_safety",
                  "S7_labour_state",
                  #"yearly_energy",
                  "nutrition_quality"
                  #"hh_comp", 
                  #"marital_status"
)

@@

mice.long$pidp <- mice.data$pidp
mice.long$weight <- mice.data$weight
mice.long$time <- mice.data$time
mice.long.colnames <- colnames(mice.long)
mice.long.colnames[which( colnames(mice.long)=="job_sec" )] <- "NSSEC"
colnames(mice.long) <- mice.long.colnames

mice_set.full <- as.mids(mice.long)

@@

```{r cross_val_handovers}
# boxplots for handovers
#.estimating from 2014-2020.
# get data for 2013- 2020
# run 2013 pop through transition probability model a few times. 
# profit. 
# run it through model a few times
# compare with real data in boxplots and ridgelines.

crossval.mice.data <- complete(mice_set.full, 1, include = F)
crossval.mice.data <- replace.missing(crossval.mice.data)
crossval.mice.data <- drop_na(crossval.mice.data)
```

```{r mice_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data[, c("pidp", "SF_12", "time")]
handover.data <- rolling.data
for (i in 1:7){
  rolling.data$SF_12 <- predict(pooled_lm, start.data)
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data)
}

@@

```{r cross_val_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data[, c("pidp", "SF_12", "time")]
handover.data <- rolling.data

for (i in 1:7){
  rolling.data$SF_12 <- predict(cross.val.sf12.model, start.data)
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data)
}

@@

lasso.min <-createTexreg(coef.names=rownames(lasso_coefs), coef = lasso.min.coefs[, "s1"])
lasso.1se <-createTexreg(coef.names=rownames(lasso_coefs), coef = lasso.1se.coefs[, "s1"])

@@

#coefficients go here.
texreg_file <- paste0(save.path, 'SF12_Lasso_coefficients.txt')
tex_label <- paste0("table: sf12_lasso_coefficients") 
tex_headers <- c(texreg::names2latex("SF_12 LASSO"))
tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "lasso"))    
lassoTexReg <-createTexreg(coef.names=rownames(lasso_coefs), coef = lasso_coefs[,1], se=lasso_coefs[,2])
#lassoTexReg2 <-createTexreg(coef.names=rownames(lasso_coefs), coef = lasso_coefs[,2])#, lasso_coefs[, 2]))#, pvalues=lasso_coefs[,2])
#matrixLassoTexReg <- matrixreg(list(lassoTexReg, lassoTexReg2))

texreg(lassoTexReg, # what model to save coefs for. 
       file=texreg_file,  # where to save.
       custom.model.names = tex_headers, # title of table.
       caption=tex_caption, # \caption command option. 
       label = tex_label, # \label command option. 
       dcolumn=T,# nice column alignment. 
       booktabs=T, # nice hlines. (recommended)
       tabular=T,# tabulr env.
       single.row = T, # nicer one row formatting.
       longtable=T, # allows tables to wrap over pages
       use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
       fontsize='small') # small font

@@

# yoink transition model function pre processing
reflect <- T
yeo_johnson <- F
include_weights <- T
depend <- "SF_12"
dependent_data <- complete(mice_set.subset)$SF_12

@@

asd <- "SF_12_last +"
glmm_model_formula_string <- "SF_12 ~ SF_12_last +
                  factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(S7_labour_state) + 
                  relevel(factor(NSSEC), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_income) + 
                  scale(I(hh_income**2)) +
                  factor(housing_quality) +
                  factor(neighbourhood_safety) +
                  factor(loneliness) + 
                  nutrition_quality + 
                  ncigs + time +
                  (1|pidp)"

@@

mice.glmm.object@beta <- -mice.glmm.object@beta 
gamma_forest_plot(mice.glmm.object, paste0(save.path, "glmm_forest.pdf"))
mice.glmm.object@beta <- -mice.glmm.object@beta

@@

start.data <- complete(glmm.mice.set, 1)
start.data <- start.data[which(start.data$time==2013),]
start.data$SF_12 <- crossval.glmm.data[which(crossval.glmm.data$time==2013),]$SF_12
start.data <- drop_na(start.data)

handover.data <- start.data

rolling.data <- start.data
for (i in 1:7){
  if (yeo_johnson) {
      rolling.data$SF_12 <- max_value - predict(yj,  predict(mice.glmm.object, start.data, type='response',allow.new.levels=T) +min_value, inverse=T)
  }
  else {
    rolling.data$SF_12 <- max_value - predict(mice.glmm.object, start.data,type='response',allow.new.levels=T) + rnorm(nrow(rolling.data), 0, 1)
  }
  #rolling.data$SF_12 <- max_value - rolling.data$SF_12
  rolling.data$SF_12 <- trunc(rolling.data$SF_12, 0 , 100)
  rolling.data$SF_12_last <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data)
}

>>>>

# baseline OLS model formula. 
raw_model_formula_string <- "SF_12 ~ factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(labour_state) + 
                  relevel(factor(NSSEC), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_income)"

@@

# get only needed variables and replace missing codes with NA.
raw.data <- update_column_names(raw.data)
raw.data <- raw.data[, append(variables, c("time", 'pidp', 'weight'))]
raw.data <- replace.missing(raw.data)

@@

# get raw data file names.
full.raw.filelist <- list.files(raw.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[21:30] # get files for 2011-2013.
# load raw data.
full.raw.data <- do.call(rbind, lapply(full.raw.filelist, read.csv))
full.raw.data <- update_column_names(full.raw.data)

# get only needed variables and replace missing codes with NA.
full.raw.data <- full.raw.data[, append(variables, c("time", 'pidp', 'weight'))]
full.raw.data <- replace.missing(full.raw.data)
```

@@

# calculate baseline OLS model.
sf12.lm <- lm(raw_model_formula_string,
              data= raw.data)
# print summary statistics
#print(summary(sf12.lm))

@@

start.data <- raw.data[which(raw.data$time==2013),]
start.data.colnames <- colnames(start.data)
start.data.colnames[which( colnames(start.data)=="job_sec" )] <- "NSSEC"
start.data.colnames[which( colnames(start.data)=="hh_netinc" )] <- "hh_income"
start.data.colnames[which( colnames(start.data)=="labour_state_raw" )] <- "labour_state"
colnames(start.data) <- start.data.colnames
handover.data <- start.data[, c("pidp", "SF_12", "time")]
rolling.data <- start.data

for (i in 1:7){
  start.data$SF_12 <- predict(sf12.lm, rolling.data)
  start.data[, 'time'] <- start.data[, 'time'] +1
  handover.data <- rbind(handover.data, start.data[, c("pidp", "SF_12", "time")])
}

@@

mice.data <- update_column_names(mice.data)

@@

mice.2018.pidp <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "pidp"]
mice.2018.time <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "time"]
mice.2018.weights <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "weight"]

lm.no.mice <- lm(mice_formula_string, data=mice.data)
forest_plot(lm.no.mice,  paste0(save.path, "no_mice_forest.pdf"))
```

@@

imp_columns <- c("SF_12",
                 'sex',
                 "NSSEC",
                 "education_state",
                 "labour_state", 
                 "region", 
                 "ethnicity", 
                 "age", 
                 "time", 
                 "hh_income",
                 "housing_quality",
                 "neighbourhood_safety",
                 "loneliness",
                 "ncigs",
                 'nutrition_quality'
)

@@

aggr_columns <- c("SF_12",
                 "NSSEC",
                 "education_state",
                 "labour_state", 
                 "ethnicity", 
                 "hh_income",
                 "neighbourhood_safety",
                 "loneliness",
                 "ncigs",
                 'nutrition_quality'
)
# initial missingness structure for 2011-2013 US data.
pdf(paste0(save.path, "total_missingness_structure1.pdf"))
aggr(mice.data[,aggr_columns], sortVars=T, prop=T, combined=F,varheight=T, oma=c(10,5,1,2), gap=0)
dev.off()

mice_columns <- c("age", 
                  "region", 
                  #"heating", 
                  "NSSEC", 
                  "ncigs",
                  "education_state",            
                  "ethnicity",
                  "loneliness",
                  "sex", 
                  "SF_12",
                  #"SF_12p",
                  #"smoker",
                  "nkids",       
                  "behind_on_bills",
                  "financial_situation",
                  "future_financial_situation",
                  "likely_move",
                  "ghq_depression",
                  "ghq_happiness",
                  "clinical_depression", 
                  "scsf1",
                  "health_limits_social",
                  #"hhsize",
                  #"housing_tenure",
                  #"urban", 
                  "housing_quality",
                  "hh_income",
                  "neighbourhood_safety",
                  "labour_state",
                  #"yearly_energy",
                  "nutrition_quality"
                  #"hh_comp", 
                  #"marital_status"
)

@@

mice.long$pidp <- mice.data$pidp
mice.long$weight <- mice.data$weight
mice.long$time <- mice.data$time
mice.long <- update_column_names(mice.long)
mice_set.full <- as.mids(mice.long)

@@

```{r mice_cross_val_handovers_data}
# boxplots for handovers
#.estimating from 2014-2020.
# get data for 2013- 2020
# run 2013 pop through transition probability model a few times. 
# profit. 
# run it through model a few times
# compare with real data in boxplots and ridgelines.

crossval.mice.data <- complete(mice_set.full, 1, include = F)
crossval.mice.data <- replace.missing(crossval.mice.data)
crossval.mice.data <- drop_na(crossval.mice.data)
crossval.mice.data <- update_column_names(crossval.mice.data)
```

```{r mice_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data#[, c("pidp", "SF_12", "time")]
handover.data <- rolling.data[, c("pidp", "SF_12", "time")]
for (i in 1:7){
  rolling.data$SF_12 <- predict(pooled_lm, newdata = rolling.data)
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data[, c("pidp", "SF_12", "time")])
}

@@

```{r cross_val_mice_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data
handover.data <- rolling.data[, c("pidp", "SF_12", "time")]

for (i in 1:7){
  rolling.data$SF_12 <- predict(cross.val.sf12.model, rolling.data)
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data[, c("pidp", "SF_12", "time")])
}

@@

lasso.min <-createTexreg(coef.names=rownames(lasso.min.coefs), coef = lasso.min.coefs[, "s1"])
lasso.1se <-createTexreg(coef.names=rownames(lasso.min.coefs), coef = lasso.1se.coefs[, "s1"])

@@

#coefficients go here.
texreg_file <- paste0(save.path, 'SF12_weighted_OLS_coefficients.txt')
tex_label <- paste0("table: sf12_weighted_coefficients") 
tex_headers <- c(texreg::names2latex("SF_12 Weighted OLS"))
tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "weighted_OLS"))

texreg(pool(sf12.lm.weighted), # what model to save coefs for. 
       file=texreg_file,  # where to save.
       custom.model.names = tex_headers, # title of table.
       caption=tex_caption, # \caption command option. 
       label = tex_label, # \label command option. 
       dcolumn=T,# nice column alignment. 
       booktabs=T, # nice hlines. (recommended)
       tabular=T,# tabulr env.
       single.row = T, # nicer one row formatting.
       longtable=T, # allows tables to wrap over pages
       use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
       fontsize='small') # small font

@@

# yoink transition model function pre processing
reflect <- T
yeo_johnson <- T
include_weights <- T
depend <- "SF_12"
dependent_data <- complete(mice_set.subset)$SF_12

@@

asd <- "SF_12_last +"
glmm_model_formula_string <- "SF_12 ~ SF_12_last +
                  factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(labour_state) + 
                  relevel(factor(NSSEC), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_income) + 
                  scale(I(hh_income**2)) +
                  factor(housing_quality) +
                  factor(neighbourhood_safety) +
                  factor(loneliness) + 
                  nutrition_quality + 
                  ncigs + time +
                  (1|pidp)"

@@

#mice.glmm.object@beta <- -mice.glmm.object@beta 
gamma_forest_plot(mice.glmm.object, paste0(save.path, "glmm_forest.pdf"))
#mice.glmm.object@beta <- -mice.glmm.object@beta

@@

start.data <- complete(glmm.mice.set, 1)
start.data <- start.data[which(start.data$time==2013),]
#start.data$SF_12 <- crossval.glmm.data[which(crossval.glmm.data$time==2013),]$SF_12
if (yeo_johnson == T) {
start.data$SF_12 <- max_value - predict(yj, start.data$SF_12 +min_value, inverse=T)
start.data <- drop_na(start.data)
} else {
  start.data$SF_12 <- max_value - start.data$SF_12
}
#start.data$SF_12_last <- max_value - start.data$SF_12_last
handover.data <- start.data

rolling.data <- start.data
for (i in 1:7){
  if (yeo_johnson == T) {
      rolling.data$SF_12 <- predict(mice.glmm.object, rolling.data, type='response',allow.new.levels=T)
      rolling.data$SF_12_last <- rolling.data$SF_12                           
      rolling.data$SF_12 <- max_value - predict(yj,  rolling.data$SF_12 +min_value+ rnorm(nrow(rolling.data), 0, 0.3), inverse=T)
  }
  else {
    rolling.data$SF_12 <- max_value - predict(mice.glmm.object, rolling.data, type='response',allow.new.levels=T) + rnorm(nrow(rolling.data), 0, 1)
    rolling.data$SF_12_last <- rolling.data$SF_12
  }
  #rolling.data$SF_12 <- max_value - rolling.data$SF_12
  rolling.data$SF_12 <- trunc(rolling.data$SF_12, 0 , 100)
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data)
}
